---
title: "KM_19.10.18_Q6"
author: "Kiana Miyamoto"
date: "10/18/2019"
output: html_document
---

```{r}
# Can you improve this analysis code?
library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug
 #trim.pdb = downsize pdb file to chainA
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")
  #create vectors with 214 residues from the s#.chainA to select for atoms from b
s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b
  #create line plots for each protein visualizing the B-factor trends and secondary structure
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```

```{r}
#create overlayed plot of all proteins(1,2,3) B-factor trends
plotb3(s1.b, sse=s1.chainA, typ="l",lwd=2, ylab="Bfactor")
points(s2.b, typ="l",col="blue", lwd=2)
points(s3.b, typ="l",col="green", lwd=1)
```

```{r}
#create a Cluster Dendrogram for quantitative comparison of B-factor trends
hc<-hclust(dist(rbind(s1.b,s2.b,s3.b)))
plot(hc)
```

Q6. How would you generalize the original code above to work with any set of input
protein structures?

```{r}
library(bio3d)
#create input vector of protein names: proteins
proteins<-c("3fxx","4AKE", "1AKE", "1E4Y")

res_vec<- function(x) {
  #read.pdb searches for name in proteins and save in s1
  s1 <- read.pdb(x)
 #trim.pdb = downsize pdb file to chainA
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
  #create vectors with 214 residues from the s#.chainA to select for atoms from b
s1.b <- s1.chainA$atom$b
  #create line plots for each protein visualizing the B-factor trends and secondary structure
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
}
#use sapply to apply res_vec over proteins to generate plots
sapply(proteins, res_vec)
```

Documentation:
1 pt - comments on what are the inputs to the function.
1 pt - what the function does and how to use it.
1 pt - what is the output of the function.

Code:
2 pt - function behaves as desired, producing the correct output and follows assignment
specifications.
2 pt - the code is efficient meaning it uses best practices such as limiting calculation
duplication.
2 pt - code is readable, meaning best practices are used including proper indentation
and whitespace used, relevant variable names, and organized in a logical manner.
1 pt - function code and call executes and is working properly
