---
title: "UKFoods"
author: "Kiana Miyamoto"
date: "10/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
UK Foods
```{r}
x <- read.csv("~/BGGN213_Bioinformatics/class08/UK_foods.csv")
```
Q1. How many rows and columns are in your new data frame named x? What R functions could you use to answer this questions?
rows: 17
columns:5
```{r}
nrow(x)
ncol(x)
dim(x)
str(x)
```
```{r}
#minus indexing
rownames(x)<- x[,1]
x<- x[,-1]
head(x)
```
```{r}
dim(x)

x <- read.csv("~/BGGN213_Bioinformatics/class08//UK_foods.csv", row.names=1)
head(x)
```
Q2. Which approach to solving the ‘row-names problem’ mentioned above do you prefer and why? Is one approach more robust than another under certain circumstances?
2nd one, because it isn't destructive code over many iterations

```{r}
barplot(as.matrix(x), beside=T, col=rainbow(nrow(x)))
```

Q3: Changing what optional argument in the above barplot() function results in the following plot?
change beside arg to F
```{r}
barplot(as.matrix(x), beside=F, col=rainbow(nrow(x)))
```

Q5: Generating all pairwise plots may help somewhat. Can you make sense of the following code and resulting figure? What does it mean if a given point lies on the diagonal for a given plot?
```{r}
pairs(x, col=rainbow(10), pch=16)
```
Q6. What is the main differences between N. Ireland and the other countries of the UK in terms of this data-set? indicated by varying points from the diagonal, however, does not specify what those are....


```{r}
# Use the prcomp() PCA function 
pca <- prcomp( t(x) )
summary(pca)
```

Q7. Complete the code below to generate a plot of PC1 vs PC2. The second line adds text labels over the data points.
```{r}
# Plot PC1 vs PC2
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1], pca$x[,2], colnames(x))
```

Q8. Customize your plot so that the colors of the country names match the colors in our UK and Ireland map and table at start of this document.
```{r}
# Plot PC1 vs PC2
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1], pca$x[,2], colnames(x), col=c("red", "orange", "blue", "green"))
abline(h=0,col="gray", lty=2)
```

```{r}
#Below we can use the square of pca$sdev , which stands for “standard deviation”, to calculate how much variation in the original data each PC accounts for.
v <- round( pca$sdev^2/sum(pca$sdev^2) * 100 )
v

## or the second row here...
z <- summary(pca)
z$importance
```

```{r}
barplot(v, xlab="Principal Component", ylab="Percent Variation")
```

```{r}
## Lets focus on PC1 as it accounts for > 90% of variance 
par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,1], las=2 )
```

Q9: Generate a similar ‘loadings plot’ for PC2. What two food groups feature prominantely and what does PC2 mainly tell us about?
```{r}

```

